When environmentally persistent pathogens transform good habitat into ecological traps

### 2016-02-17 Manuscript/manuscript.tex
Edits to discussion

The framework used here provides general insights, but generating predictions for specific systems requires a careful consideration of the model's simplifying assumptions.
In particular, the stochastic patch occupancy model, though well suited for capturing realistic spatial structure, assumes that within-patch dynamics (i.e. growth to carrying capacity and local pathogen spread) are rapid relative to among-patch dynamics (i.e. colonization and extinction).
This is often a reasonable assumption, especially for relatively small populations in a highly fragmented landscape \cite{Hanski2003}, but may not capture the non-linearities and nuances of host-pathogen dynamics in a single population.
In addition, we assume that patch quality only influences population size and the resulting colonization and extinction processes, but quality could also influence pathogen dynamics within a patch through effects on host density or condition \cite{Becker2015}.
Patch quality may also influence the size or infectiousness of the environmental reservoir, either directly, by influencing the decay rate of the pathogen in the environment, or indirectly, through the effects on population size and overall shedding rate.
More generally, the dynamics and establishment of environmental reservoirs could be influenced by a number of other factors, such as the length of infectious occupancy or the occupancy history of a patch.
The potential importance of these additional mechanisms remains an open question for future work and needs to be considered before applying this framework to any given system of interest.
Fortunately, the SPOM framework is fairly flexible and future work could extend the model presented here to include additional processes, e.g. by making transmission parameters also function of patch quality.

>>>>

In addition to helping to guide the prioritization of conservation efforts, our model may also help to inform the deployment of disease control measures, including the treatment of the environmental reservoir itself.
In the case of African herbivores and their tick-borne pathogens, for example, the environmental reservoir created by ticks and livestock hosts could be controlled through the treatment of cattle with acaricides \cite{Keesing2013}.
Our work suggests that such efforts would likely be most effective if focused on cattle that share high quality habitat with the wildlife hosts of interest.

The framework used here provides general insights, but generating predictions for specific systems requires a careful consideration of the model's simplifying assumptions.
In particular, the stochastic patch occupancy model, though well suited for capturing realistic spatial structure, does not account for within-patch dynamics.
As a result, patches are modeled as either unoccupied or occupied by a host population with a fixed size and discrete disease status -- either susceptible or infected.
This occupancy framework assumes that within-patch dynamics (i.e. growth to carrying capacity and local pathogen spread) are rapid relative to among-patch dynamics (i.e. colonization and extinction).
Moreover, it assumes that the pathogen spreads widely within a patch and affects host dynamics primarily by reducing population size.
These are often reasonable assumptions, especially for relatively small populations in a highly fragmented landscape \cite{Hanski2003}, but may not capture the non-linearities and nuances of host-pathogen dynamics in a single population.

In addition, we assume that patch quality only influences population size and the resulting colonization and extinction processes, but quality could also influence pathogen dynamics within a patch through effects on host density or condition \cite{Becker2015}.
Patch quality may also influence the size or infectiousness of the environmental reservoir, either directly, by influencing the decay rate of the pathogen in the environment, or indirectly, through the effects on population size and overall shedding rate.
More generally, the dynamics and establishment of environmental reservoirs could be influenced by a number of other factors, such as the length of infectious occupancy or the occupancy history of a patch.
The potential importance of these additional mechanisms remains an open question for future work and needs to be considered before applying this framework to any given system of interest.
Fortunately, the SPOM framework is fairly flexible and future work could extend the model presented here to include additional processes, e.g. by making transmission parameters also function of patch quality.
